\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-25 paper_resources/doc/paper.tex
Some polishing

\fxnote{Dati kompleksnost velicine skupa perturbacija u odnosu na velicinu drveta i dopustene promjene. ACA: tesko je ovo izracunati, u principu za svaki cvor, dakle, celo drvo ili bilo koje pod-drvo ide ovaj spisak gore, dakle, neki red velicine d*C(T), gde je C(T) broj poddrveca iz datog drveta T, a d je neka konstanta, tj. broj mogucih perturbacija jednog poddrveta. Sad koliko ima poddrveca, moze se izracunati ovako nekako, ali ne bih to ukljucivao osim ako nam ne zatraze. Dakle, za drvo T broj poddrveca $C(T) = C(T.left)+C(T.right)+1$ (ovo jedan za bas to celo drvo u datom rekurzivnom koraku), sad kad je drvo T od N cvorova balansirano, dobija se diferencna jednacina $C(N)=2*C(N/2)+1$, a njeno resenje je O(N) mislim. Kad se raspise dobija se $C(N) = 2*2*C(N/4)+2 = 2*2*2*C(N/8)+3 = ... = 2^(\log_2(N))*C(1)+\log_2(N) = N + \log_2(N) = O(N)$ posto je $C(1)=1$, to je list drveta}

@@

Our \textsc{RILS}--\textsc{ROLS} algorithm is implemented in Python version 3.x. The experiments concerning our method are performed on a machine xx with xx GHz and a memory limit of xx GB in single-threaded mode.  
The following 14 algorithms are compared to our approach: \textsc{AI-Feynman}, \textsc{GOMEA}, \textsc{Afp-Fe}, \textsc{Itea}, \textsc{Afp}, \textsc{Dsr}, \textsc{Operon}, the \textsc{Gplearn} from python package \textsc{gplearn}~\cite{stephens2016genetic}, \textsc{Sbp-Gp}, \textsc{Eplex}, \textsc{Bsr}, \textsc{Feat}, \textsc{Ffx}, \textsc{Mrgp}.

For the (14) competitors, the results are exported from \url{https://cavalab.org/srbench/results/}, also reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1h, i.e., 3600 seconds; note that the same runtime limit was used for all (14) competitors, see~\cite{la2021contemporary}.

All (15) competitor algorithms were run ten times per each instance.

@@

SRBench is an open-source benchmarking project which merge a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and an environment for analysis, see~\cite{la2021contemporary}. It brought researchers an easy handled environment to evaluation, comparisons, and analysis of their methods. Among all benchmark sets, we involve two real-world inspired benchmark sets, the ground-truth regression problems, into our computational evaluation: 
\begin{itemize}
	\item \textsc{Feynman} benchmark set, inspired from physics and formulas/models that correspond to various natural laws.  
	It consists of $10^5$  samples, each described in~\cite{udrescu2020ai}. Some exact models (equations) of problem instances from this benchmark are shown in Table~\ref{tab:Feynamn-Eq}.  Among them, 116 regression problem instances are chosen in the experimental evaluation, with accordance to~\cite{la2021contemporary}.

@@

	\item \textsc{Strogatz} benchmark set from~\cite{la2016inference}. 
	Each dataset is represented by a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict rate of change of the subsequent state w.r.t\  the current two states on which it depends. These equations describes natural processes described by non-linear dynamics  exhibiting chaos.  The equations for some of the datasets that belong to this benchmark are shown in Table~\ref{table:strogatz-ODEs}. These problem instances are downloadable from \url{https://github.com/lacava/ode-strogatz}. In overall, 14 ground-truth models are used to generate the problem instances used in the comparison, see~\cite{la2021contemporary}.

@@

\end{itemize}
The above-mentioned benchmark sets may be biased as models which describe physical laws that usually impose   symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be, in-advance, significantly reduced. Thus, some methods may get significant benefit with this regard, while some others may be in the subordinate situation. 
In order to permit any kind of favoritism in the input data, we generate a set of randomly generated (non-biased) instances, labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

@@

In future work, one could think of constructing a hybrid of \textsc{RILS-ROLS} method with some other meta-heuristics to further boost the quality of the obtained results. For example, replacing ILS scheme of \textsc{RILS-RILS} with a more general Variable neighbourhood search (VNS) scheme is a reasonable option to give it a try. VNS   utilizes a more stronger and systematic system of solutions diversification and thus escaping from local optima could be much easier.  Additionally, as proof of the efficiency and benefits of our algorithm, this tool could be applied to solving problems from various problem domains out of the area of computer science, for example from physics or chemistry, to help for setting up hypotheses or to obtain valuable insights on experimental evaluations.

>>>>

%\fxnote{Dati kompleksnost velicine skupa perturbacija u odnosu na velicinu drveta i dopustene promjene. ACA: tesko je ovo izracunati, u principu za svaki cvor, dakle, celo drvo ili bilo koje pod-drvo ide ovaj spisak gore, dakle, neki red velicine d*C(T), gde je C(T) broj poddrveca iz datog drveta T, a d je neka konstanta, tj. broj mogucih perturbacija jednog poddrveta. Sad koliko ima poddrveca, moze se izracunati ovako nekako, ali ne bih to ukljucivao osim ako nam ne zatraze. Dakle, za drvo T broj poddrveca $C(T) = C(T.left)+C(T.right)+1$ (ovo jedan za bas to celo drvo u datom rekurzivnom koraku), sad kad je drvo T od N cvorova balansirano, dobija se diferencna jednacina $C(N)=2*C(N/2)+1$, a njeno resenje je O(N) mislim. Kad se raspise dobija se $C(N) = 2*2*C(N/4)+2 = 2*2*2*C(N/8)+3 = ... = 2^(\log_2(N))*C(1)+\log_2(N) = N + \log_2(N) = O(N)$ posto je $C(1)=1$, to je list drveta}

@@

Our \textsc{RILS}--\textsc{ROLS} algorithm is implemented in Python version 3.x. The experiments concerning our method are performed on a machine xx with xx GHz and a memory limit of xx GB in single-threaded mode.  
The following 13 algorithms are compared to our approach: \textsc{AI-Feynman}, \textsc{GOMEA}, \textsc{Afp-Fe}, \textsc{Itea}, \textsc{Afp}, \textsc{Dsr}, \textsc{Operon}, the \textsc{Gplearn} from python package \textsc{gplearn}~\cite{stephens2016genetic}, \textsc{Sbp-Gp}, \textsc{Eplex}, \textsc{Bsr}, \textsc{Feat}, \textsc{Ffx}, \textsc{Mrgp}.

For the (13) competitors, the results are exported from \url{https://cavalab.org/srbench/results/}, reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1h, i.e., 3600 seconds; note that the same runtime limit was used in case of all (13) competitors, see~\cite{la2021contemporary}.

All    algorithms were run ten times per each problem instance.

@@

SRBench is an open-source benchmarking project which merge a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and an environment for analysis, see~\cite{la2021contemporary}. It brings researchers an easy handled environment to evaluate, compare, and analyze their methods. Among all benchmark sets, we involve two real-world inspired ground-truth benchmark sets that represent types of regression problems, into our computational evaluation: 
\begin{itemize}
	\item \textsc{Feynman} benchmark set, inspired from physics and formulas/models that describes   various natural laws.  
	It consists of $10^5$  samples, each described in~\cite{udrescu2020ai}. Some exact models (equations) of problem instances from this benchmark are shown in Table~\ref{tab:Feynamn-Eq}.  Among them, 116 regression problem instances are chosen in the experimental evaluation, with accordance to the results reported in~\cite{la2021contemporary}.

@@

	\item \textsc{Strogatz} benchmark set from~\cite{la2016inference}. 
	Each dataset is represented by a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict rate of change of the subsequent state w.r.t\  in dependence of the current two states. These equations describes natural processes described by non-linear dynamics  exhibiting chaos.  The equations for some of the datasets that belong to this benchmark are shown in Table~\ref{table:strogatz-ODEs}. These problem instances are downloadable from \url{https://github.com/lacava/ode-strogatz}. In overall, 14 ground-truth models are used to generate the problem instances used in the comparison, also in accordance with~\cite{la2021contemporary}.

@@

\end{itemize}
The above-mentioned benchmark sets may be biased as models which describe physical laws that usually impose   symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be, in-advance, significantly reduced. Thus, some methods may get significant benefit with this regard, while some others may be in the subordinate situation. 
In order to permit any kind of favoritism in the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

@@

In future work, one could think of constructing a hybrid of \textsc{RILS-ROLS} method with some other meta-heuristics to further boost the quality of the obtained results. For example, replacing ILS scheme of \textsc{RILS-RILS} with a more general Variable neighbourhood search (VNS) scheme is a reasonable option to give it a try. VNS   utilizes a more stronger and systematic system of solutions diversification and thus escaping from local optima could be much easier and the convergence much faster.  Additionally, as proof of the efficiency and benefits of our algorithm, this tool could be applied to solving problems from various problem domains out of the area of computer science, for example from physics or chemistry, to help for setting up reasonable hypotheses or to obtain valuable insights on experimental evaluations.
